---
title: "Final"
author: "Xiaotai Chai"
date: "5/9/2018"
output:
  html_document:
    self_contained: no
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("dslabs")
#install.packages("devtools")
#install_github("ropensci/plotly")
#devtools::install_github("dgrtwo/gganimate")
#install.packages('animation')
#devtools::install_github("yihui/animation")
#devtools::install_github('tidyverse/ggplot2')

library(ggplot2)
library(animation)
library(gganimate)
library(devtools)
library(tidyverse)
library(dslabs)
library(plotly)

# knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.show = "animate")

diseases <- us_contagious_diseases %>% filter(disease != "Polio")
```

### How are overall disease trends changing over time? Are some diseases more of a threat today than others?
*Tips: Plotly graphs are interactive. Click on legend entries to toggle traces, click-and-drag on the chart to zoom, double-click to autoscale, shift-and-drag to pan.*
```{r message=FALSE, warning=FALSE}
# add a per 10,000 people rate column, and removes Alaska and Hawaii since they only became states in the late 50s.
diseases.rate <- diseases %>%
  filter(!state%in%c("Hawaii","Alaska")) %>%
  mutate(rate = (count / weeks_reporting) * 52 / (population / 10000))
  
p1 <- ggplot(aes(year, rate, color = disease), data= diseases.rate) +  
  geom_point(size = 1, alpha= 0.5) +
  geom_smooth(method = "loess") +
  scale_y_continuous(limits = c(0,250)) +
  labs(title = "Disease Trends Over Time", x = "Year",
    y = "Disease Rate (per 10,000 people)", color = "Disease") + 
  theme(legend.position = c(.95, .95),legend.justification = c("right", "top"),legend.box.just = "right",legend.margin = margin(6, 6, 6, 6))

ggplotly(p1, width = 1000, height = 600)

```
From the plot, we can see that Measles was a serious and over spread disease before 1968 and less and less people got this disease starting 1970s. The trends for most diseases are decreasing. Only for Pertussis, it decreased a lot during 1960s but started to increase a little bit from 1970s and become a threat comparing to others. Adding facet_wrap to separate each disease makes the trend of each disease clearer.

```{r message=FALSE, warning=FALSE}
p2 <- ggplot(aes(year, rate, color = disease), data= diseases.rate) +  
  geom_point(size = 1, alpha= 0.5) +
  geom_smooth(method = "loess", color = "black") +
  scale_y_continuous(limits = c(0,30)) +
  facet_wrap(~ disease) +
  labs(title = "Disease Trends Over Time", x = "Year",
    y = "Disease Rate (per 10,000 people)", color = "Disease") + 
  theme(legend.position = "none")

ggplotly(p2, width = 1000, height = 600)
```

### Using the murder count and disease count, along with population, do some states or regions seem more dangerous than others (as of 2010)? Find a way to visualize overall “risk” for each state in 2010.
The plot shows that in 2010, Lowa and New York had an overall high risk of living. Lowa had a high risk of getting Pertussis and New York had a high risk of getting Hepatitis A. Alos, District of Columbia had a high risk of geting murdered.
```{r message=FALSE, warning=FALSE}
diseases.2010 <- diseases %>%
  filter(year == "2010") %>%
  mutate(rate = (count / weeks_reporting) * 52 / (population / 10000)) %>% 
  select(disease, state, rate)

murders.2010 <- murders %>% 
  mutate(rate = total / (population / 10000), disease = "Murder") %>% 
  select(state, disease, rate)

data.2010 <- rbind(diseases.2010, murders.2010)

p3 <- ggplot() +
  geom_bar(aes(y = rate, x = state, fill = disease), data = data.2010, stat="identity")+
  theme(axis.text = element_text(angle=90))+
  labs(title = "Diseases and Murders on States 2010", x = "State", y = "Disease/Murder Rate (per 10,000 people)", fill = "Disease")

ggplotly(p3, width = 1000, height = 600)
```

### Is there a geographic pattern to illnesses? After accounting for population, are some states or regions more at risk for certain diseases, or is the risk fairly constant in all locations?
*The attached gif plot will show you the rate of diseases on each state for each year.*
```{r message=FALSE, warning=FALSE, fig.width=16,fig.height=10}
# include Alaska and Hawaii since we want the location information
diseases.location <- diseases %>%
  mutate(rate = (count / weeks_reporting) * 52 / (population / 10000))
p4 <- ggplot(diseases.location, aes(state, rate, color = disease, frame = year)) +
  geom_point(size = 3)+
  scale_y_continuous(limits = c(0,170)) +
  theme(axis.text = element_text(size = 14, angle=90),axis.title = element_text(size = rel(2)), plot.title = element_text(size = rel(2)), legend.text = element_text(size = 12), legend.title = element_text(size = rel(2)))+
  labs(title = "Diseases on States", x = "State", y = "Disease Rate (per 10,000 people)", color = "Disease")

p5 <- ggplot(diseases.location, aes(state, rate, color = disease, frame = year)) +
  geom_point(size = 2)+
  scale_y_continuous(limits = c(0,170)) +
  theme(axis.text = element_text(size = 14, angle=90),axis.title = element_text(size = rel(2)), plot.title = element_text(face = "bold",size = rel(2.5)), legend.text = element_text(size = 12), legend.title = element_text(size = rel(2)))+
  labs(title = "Diseases on States", x = "State", y = "Disease Rate (per 10,000 people)", color = "Disease")
p5
#gganimate(p4, "p4.gif")
```
